BPMN Analyzer 2.0: Instantaneous, Comprehensible, and Fixable Control Flow Analysis for Realistic BPMN Models

### 2024-08-01 paper.tex
Cleanup and typos

%% hf: enable header and footer.
\documentclass[
onecolumn,
]{ceurart}

% One can fix some overfills
% \sloppy

@@

% Rights management information. CC-BY is the default license.
\copyrightyear{2024}
\copyrightclause{Copyright for this paper by its authors.
  Use permitted under Creative Commons License Attribution 4.0
  International (CC BY 4.0).}

@@

\author[1]{Tim Kräuter}
[email=tkra@hvl.no]
\author[1]{Patrick Stünkel}
[email=past@hvl.no] % patrick.stuenkel@hvl.no
\author[1]{Adrian Rutle}
[email=aru@hvl.no]
\author[1]{Yngve Lamo}
[email=yla@hvl.no]
\author[2,1]{Harald König}
[email=harald.koenig@fhdw.de]
\address[1]{Western Norway University of Applied Sciences, Bergen, Norway}
\address[2]{FHDW Hannover, Germany}

@@

% Solution
In this paper, we describe a new tool, the \textit{BPMN Analyzer 2.0}\footnote{
In the following, we will use BPMN Analyzer to refer to the BPMN Analyzer 2.0.}, for analyzing BPMN process models to detect control flow errors \textit{already} during modeling.
\autoref{fig:overview} shows an overview of the tool.
The UI is based on the popular \textit{bpmn.io} ecosystem, while the analysis is implemented in \textit{Rust} for optimal performance and memory efficiency.
We perform a breadth-first state space exploration during which we check soundness and safeness~\cite{corradiniClassificationBPMNCollaborations2018} \textit{on the fly} to uncover control flow errors.
Consequently, the tool can detect deadlocks, livelocks, starvation, dead activities, and lack of synchronization in BPMN models.
The BPMN Analyzer is open source and accessible online alongside a video demonstration\footnote{Tool: \url{https://timkraeuter.com/bpmn-analyzer-js/}, Video: \url{https://www.youtube.com/watch?v=Nv2W-hXNZYA}}~\cite{krauterInstantaneousComprehensibleFixable2024}.

@@

% Why is it significant for the BPM field?
Fahland et al.~\cite{fahlandAnalysisDemandInstantaneous2011} describe \textit{coverage}, \textit{immediacy}, and \textit{consumability} as the main challenges for users unaccustomed to formal analysis.
The BPMN Analyzer addresses all these challenges since it supports the most common BPMN elements used in practice (coverage), provides \textit{instantaneous} results (immediacy), and a \textit{comprehensible} user interface (consumability), even including suggestions for fixes.
Developers of industrial BPMN software also like our tool, especially the End-to-end user journey~\cite{krauterInstantaneousComprehensibleFixable2024}.
Thus, this supports our claim that the UI is easy for users unfamiliar with formal analysis to understand.

@@

% Must have section according to the description
% A section discussing the innovations of the tool or resource to the BPM community and its main characteristics or features
\section{Innovations} \label{sec:innovations} % Main tool description and showing the innovations
The BPMN Analyzer has three main innovations: \textbf{instantaneous}, \textbf{comprehensible}, and \textbf{fixable} control flow error detection.
In this section, we will present the innovations, and more details can be found in our extended paper~\cite{krauterInstantaneousComprehensibleFixable2024}.

@@

\begin{table}[ht]
\scalebox{1}{ % Somehow needed
    \begin{minipage}{.5\linewidth}
	   \centering
	   \caption{Benchmark results of the parallel branches models}
	   \label{tab:parallel-branches-benchmark}
	   \SetTblrInner{colsep=2pt}
	   \begin{tblr}{
	   		column{1-X} = {c},
	   		column{Y-Z} = {r},
	   		hline{1, 2, Z} = {-}{1.2pt, solid}, % Z is the last row/column
	   		hline{8, 9} = {-}{dashed},
	   		vline{2-Y} = {2-Z}{solid}, % Y is the second to last row/column
	   	}
	       \textbf{Branches} & \textbf{Branch Length} &\textbf{Runtime} & \textbf{States} \\
	       5 & 1 & 1 ms& 35 \\
	       10 & 1 & 3 ms& 1.027 \\
	       15 & 1 & 161 ms& 32.771 \\
	       16 & 1 & 360 ms& 65.539 \\
	       17 & 1 & 790 ms& 131.075 \\
	       20 & 1 & 8.803 ms& 1.048.579 \\
	       5 & 5 & 14 ms& 7.779 \\
	       3 & 20 & 11 ms& 9.264 \\
	   \end{tblr}
    \end{minipage}
    % Add extra space in the middle
    \begin{minipage}{.5\linewidth}
	   \centering
	   \caption{Benchmark results of the realistic BPMN models}
	   \label{tab:realistic-models-benchmark}
	   \SetTblrInner{colsep=2pt}
	   \begin{tblr}{
	   		column{Y-Z} = {r},
	   		hline{1, 2, Z} = {-}{1.2pt, solid}, % Z is the last row/column
	   		hline{5} = {-}{dashed},
	   		vline{2-Y} = {2-Z}{solid}, % Y is the second to last row/column
	   	}
	   	   \textbf{Model name} &\textbf{Runtime} & \textbf{States} \\
	   	   e001~\cite{houhouFirstOrderLogicVerification2022} & 1 ms & 39 \\
	   	   e002~\cite{houhouFirstOrderLogicVerification2022} & 1 ms & 39 \\
	   	   e020~\cite{houhouFirstOrderLogicVerification2022} & 10 ms & 5356 \\
	   	   credit-scoring-async\footref{footnote:camundaResearch} & 1 ms & 60 \\
	   	   credit-scoring-sync\footref{footnote:camundaResearch} & 1 ms & 140 \\
	   	   dispatch-of-goods\footref{footnote:camundaResearch} & 1 ms & 103\\
	   	   recourse\footref{footnote:camundaResearch} & 1 ms & 77 \\
	   	   self-service-restaurant\footref{footnote:camundaResearch} & 1 ms & 190 \\
	   \end{tblr}
    \end{minipage} 
}
\end{table}

@@

\section{Maturity of the Tool} \label{sec:maturity}
The BPMN Analyzer is incorporates many findings from our previous work~\cite{krauterHigherorderTransformationApproach2024} while focusing on instantaneous and understandable error detection, as described in the previous section.
The tool is open source~\cite{krauterInstantaneousComprehensibleFixable2024}, and we ensure high code quality by employing industry best practices such as rigorous static analysis and testing.
Furthermore, we received positive feedback from companies in the BPMN process orchestration space~\cite{krauterInstantaneousComprehensibleFixable2024}.

>>>>

\documentclass[onecolumn]{ceurart}

@@

% Rights management information.
\copyrightyear{2024}
\copyrightclause{Copyright for this paper by its authors.
  Use permitted under Creative Commons License Attribution 4.0
  International (CC BY 4.0).}

@@

\author[1]{Tim Kräuter}
[email=tkra@hvl.no]
\author[1]{Patrick Stünkel}
[email=past@hvl.no]
\author[1]{Adrian Rutle}
[email=aru@hvl.no]
\author[1]{Yngve Lamo}
[email=yla@hvl.no]
\author[2,1]{Harald König}
[email=harald.koenig@fhdw.de]
\address[1]{Western Norway University of Applied Sciences, Bergen, Norway}
\address[2]{FHDW Hannover, Germany}

@@

% Solution
In this paper, we describe a new tool, the \textit{BPMN Analyzer 2.0}\footnote{
In the following, we will use BPMN Analyzer to refer to the BPMN Analyzer 2.0.}, for analyzing BPMN process models to detect control flow errors \textit{already} during modeling.
\autoref{fig:overview} shows an overview of the tool.
The UI is based on the popular \textit{bpmn.io} ecosystem, while the analysis is implemented in \textit{Rust} for optimal performance and memory efficiency.
We perform a breadth-first state space exploration to check soundness and safeness~\cite{corradiniClassificationBPMNCollaborations2018} \textit{on the fly} to uncover control flow errors.
Consequently, the tool can detect deadlocks, livelocks, starvation, dead activities, and lack of synchronization in BPMN models.
The BPMN Analyzer is open source and accessible online alongside a video demonstration\footnote{Tool: \url{https://timkraeuter.com/bpmn-analyzer-js/}, Video: \url{https://www.youtube.com/watch?v=Nv2W-hXNZYA}}~\cite{krauterInstantaneousComprehensibleFixable2024}.

@@

% Why is it significant for the BPM field?
Fahland et al.~\cite{fahlandAnalysisDemandInstantaneous2011} describe \textit{coverage}, \textit{immediacy}, and \textit{consumability} as the main challenges for users unaccustomed to formal analysis.
The BPMN Analyzer addresses all these challenges since it supports the most common BPMN elements used in practice (coverage), provides \textit{instantaneous} results (immediacy), and a \textit{comprehensible} user interface (consumability), even including suggestions for fixes.
Developers of industrial BPMN software also like our tool, especially the End-to-end user journey~\cite{krauterInstantaneousComprehensibleFixable2024}.
Thus, this supports our claim that the UI is understandable for users unfamiliar with formal analysis.

@@

\section{Innovations} \label{sec:innovations} % Main tool description and showing the innovations
The BPMN Analyzer has three main innovations: \textbf{instantaneous}, \textbf{comprehensible}, and \textbf{fixable} control flow error detection.
In this section, we will present the innovations, and more details can be found in our extended paper~\cite{krauterInstantaneousComprehensibleFixable2024}.

@@

\begin{table}[ht]
\scalebox{1}{ % Somehow needed
    \begin{minipage}{.5\linewidth}
	   \centering
	   \caption{Benchmark results of the parallel branches models}
	   \label{tab:parallel-branches-benchmark}
	   \SetTblrInner{colsep=2pt}
	   \begin{tblr}{
	   		column{1-X} = {c},
	   		column{Y-Z} = {r},
	   		hline{1, 2, Z} = {-}{1.2pt, solid}, % Z is the last row/column
	   		hline{8, 9} = {-}{dashed},
	   		vline{2-Y} = {2-Z}{solid}, % Y is the second to last row/column
	   	}
	       \textbf{Branches} & \textbf{Branch Length} &\textbf{Runtime} & \textbf{States} \\
	       5 & 1 & 1 ms& 35 \\
	       10 & 1 & 3 ms& 1.027 \\
	       15 & 1 & 161 ms& 32.771 \\
	       16 & 1 & 360 ms& 65.539 \\
	       17 & 1 & 790 ms& 131.075 \\
	       20 & 1 & 8.803 ms& 1.048.579 \\
	       5 & 5 & 14 ms& 7.779 \\
	       3 & 20 & 11 ms& 9.264 \\
	   \end{tblr}
    \end{minipage}
    \begin{minipage}{.5\linewidth}
	   \centering
	   \caption{Benchmark results of the realistic BPMN models}
	   \label{tab:realistic-models-benchmark}
	   \SetTblrInner{colsep=2pt}
	   \begin{tblr}{
	   		column{Y-Z} = {r},
	   		hline{1, 2, Z} = {-}{1.2pt, solid}, % Z is the last row/column
	   		hline{5} = {-}{dashed},
	   		vline{2-Y} = {2-Z}{solid}, % Y is the second to last row/column
	   	}
	   	   \textbf{Model name} &\textbf{Runtime} & \textbf{States} \\
	   	   e001~\cite{houhouFirstOrderLogicVerification2022} & 1 ms & 39 \\
	   	   e002~\cite{houhouFirstOrderLogicVerification2022} & 1 ms & 39 \\
	   	   e020~\cite{houhouFirstOrderLogicVerification2022} & 10 ms & 5356 \\
	   	   credit-scoring-async\footref{footnote:camundaResearch} & 1 ms & 60 \\
	   	   credit-scoring-sync\footref{footnote:camundaResearch} & 1 ms & 140 \\
	   	   dispatch-of-goods\footref{footnote:camundaResearch} & 1 ms & 103\\
	   	   recourse\footref{footnote:camundaResearch} & 1 ms & 77 \\
	   	   self-service-restaurant\footref{footnote:camundaResearch} & 1 ms & 190 \\
	   \end{tblr}
    \end{minipage} 
}
\end{table}

@@

\section{Maturity of the Tool} \label{sec:maturity}
The BPMN Analyzer incorporates many findings from our previous work~\cite{krauterHigherorderTransformationApproach2024} while focusing on instantaneous and understandable error detection, as described in the previous section.
The tool is open source~\cite{krauterInstantaneousComprehensibleFixable2024}, and we ensure high code quality by employing industry best practices such as rigorous static analysis and testing.
Furthermore, we received positive feedback from companies in the BPMN process orchestration space~\cite{krauterInstantaneousComprehensibleFixable2024}.
